%&pdflatex
\addcontentsline{toc}{chapter}{Acknowledgements}
% \addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics
\begin{center}
  \large Acknowledgements
\end{center}

My thanks go to Prof. P. Le Tallec, dear supervisor of the thesis. I thank him for having assured the scientific quality in this work and for having also entrusted to me the responsibilities of my scientific productions. I will not repeat what was said on the day of the defense, but thanks to him I discovered France and polytechnique, a landscape that I once know little and far away. I could not have imagined having a better advisor and mentor for my Ph.D. studies.

I would like to thank Prof. H. Maitournam for helping me, at the crucial time of my thesis, to analyze the problems presented in this work.

I would then like to thank all the members of the jury for their work.

%I would also like to thank the two rapporteurs: Mr Combescure of the INSA of Lyon and Mr Maurini of Paris VI. I thank them for having generally appreciated the manuscript and for having formulated the appropriate scientific remarks. 

I take this opportunity to say thanks to Mme.Ida Haoult for her advice on the technical support and experiments issued by CETIM.

This thesis would not have been possible without the generous support from the Chaire PSA and the Laboratoire de Mécanique des Solides of Ecole Polytechnique to whom I express my profound gratitude.

	I would like to thank my parents for all their love and
	encouragement that have overcome the barriers of the time difference and
	geographical distance for the past three years.

Finally, I would like to thank all those whom I met and whom I did not have a chance to say thank you.