%&pdflatex
\addcontentsline{toc}{chapter}{Acknowledgements}
% \addtocontents{toc}{\vspace{1em}}  % Add a gap in the Contents, for aesthetics

\begin{center}
  \large Acknowledgements
  
\end{center}
My thanks go to Prof. P. Le Tallec, dear supervisor of the thesis. I thank him for having assured the scientific quality in this work and for having also entrusted to me the responsibilities of my scientific productions. I will not repeat what was said on the day of the defense, but thanks to him I discovered France and polytechnique, a landscape that I once know little and far away. I could not have
imagined having a better advisor and mentor for my Ph.D. studies.

I would then like to thank all the members of the jury for their work. I thank Mr Damamme for having moved from Gramat to chair my thesis. I would also like to express my appreciation to Ms. de Lorenzis for physically participating in the jury at a distance of approximately \ SI {900} {km} from her place of work. Thank you also to Mr. Frankfort who has agreed to consider this thesis.

I would like to thank Prof. H. Maitournam for helping me, at the crucial time of my thesis, to analyze the problems presented in this work.

I would also like to thank the two rapporteurs: Mr Combescure of the INSA of Lyon and Mr Maurini of Paris VI. I thank them for having generally appreciated the manuscript and for having formulated the appropriate scientific remarks. I take this opportunity to say thanks to Mme.Ida H


This thesis would not have been possible without the generous support from the
Chaire PSA and the Laboratoire de Mécanique des Solides of Ecole Polytechnique to whom I express my profound gratitude.

	I would like to thank my parents for all their love and
	encouragement that have overcome the barriers of the time difference and
	geographical distance for the past five years.


Finally, I would like to thank all those whom I met and whom I did not have a chance to say thank you.